---
title: "ev_prediction"
author: "Hir Brahmbhatt"
date: "2024-04-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Data Exploration File

```{r}
library(caTools)
set.seed(123)
ev_data = read.csv("ev_adoption_data.csv")
head(ev_data)

split = sample.split(ev_data$total_ev_count, SplitRatio = 0.8)

# training data
ev_train = subset(ev_data, split==TRUE)

# testing data
ev_test = subset(ev_data, split == FALSE)

#correlation matrix
cor(subset(ev_train, select = -c(ZIPCODE,AVG_HHSZ_OWNER,AVG_HHSZ_RENTER,TP_SEX_RATIO, TP_MED_AGE,TP_18Y_OVR_SXRATIO, TP_65Y_OVR_SXRATIO) ))


```




```{r}

#multiple linear regression to investigate the relationship between the total number of electric vehicle (EV) registrations (total_ev_count) within each ZIP code and various other demographic and socioeconomic variables. Excluded variables with insufficient data and the individual "ZIPCODE" column This model aims to identify which factors are significant predictors of EV adoption rates across different ZIP codes.
lm_model <- lm(total_ev_count ~ . - AVG_HHSZ_OWNER - AVG_HHSZ_RENTER - TP_SEX_RATIO - TP_MED_AGE - TP_18Y_OVR_SXRATIO - TP_65Y_OVR_SXRATIO - ZIPCODE, data = ev_train)
summary(lm_model)
```
From this result, we can see that OCCU_NAT(p = 0.000857), INDUS_AGRI(p = 1.42e-05), OWNER_OCC(p = 3.95e-12), TP_62Y_OVR(p = 2.19e-05) have highest significant level. 
GRAD_PRO(p = 0.002636 ) and TWO_VEHICLES(p= 0.001610) have second highest significant level.
OCCU_SEV, OCCU_SA_OF, BACHE_DEG , TP_35Y_44Y, and TP_55Y_59Y have the third highest significant level.

Except for GRAD_PRO(Graduate or professional degree) and TWO_VEHICLES(2 vehicles available) variables that have a positive impact on the EV_Adoption rate, all the other variables demonstrate significant levels having negative impact on the EV Adoption rate.

Overall, we can see that the Occupation Category has the highest impact on general EV adoption rate(cuz it has highest # of variables that show significant difference.)




```{r}
#residual plot
plot(lm_model)
summary(residuals(lm_auto))
```
It seems we do have non-linear associations(the shape looks heteroscedasticity) and outliers.




Try to transform variables to squareroot form.
```{r}
# List of variables to exclude (including response variable if it's part of ev_train)
variables_to_exclude <- c("AVG_HHSZ_OWNER", "AVG_HHSZ_RENTER", "TP_SEX_RATIO", 
                          "TP_MED_AGE", "TP_18Y_OVR_SXRATIO", "TP_65Y_OVR_SXRATIO", "ZIPCODE", "total_ev_count")

# Create a subset of ev_train without the variables to exclude
ev_train_subset <- ev_train[, !(names(ev_train) %in% variables_to_exclude)]

# Apply the square root transformation to the independent variables in the subset
ev_train_subset_sqrt <- data.frame(lapply(ev_train_subset, sqrt))

# Add the response variable back in after transformations
ev_train_subset_sqrt$total_ev_count <- ev_train$total_ev_count


lm_model_sqrt <- lm(total_ev_count ~ ., data = ev_train_subset_sqrt)


summary(lm_model_sqrt)


```

Try to transform variables to log form.
```{r}
# List of variables to exclude (including response variable if it's part of ev_train)
variables_to_exclude <- c("AVG_HHSZ_OWNER", "AVG_HHSZ_RENTER", "TP_SEX_RATIO", 
                          "TP_MED_AGE", "TP_18Y_OVR_SXRATIO", "TP_65Y_OVR_SXRATIO", "ZIPCODE", "total_ev_count")

# Create a subset of ev_train without the variables to exclude
ev_train_subset <- ev_train[, !(names(ev_train) %in% variables_to_exclude)]

# Before applying log, make sure there are no zero or negative values as they are undefined for log transformation
# Assuming all values in ev_train_subset are positive. If not, you need to add a small constant to avoid log(0)
# For example: ev_train_subset + 1

# Apply the log transformation to the independent variables in the subset
ev_train_subset_log <- data.frame(lapply(ev_train_subset, function(x) log(x + 1)))  # Adding 1 to avoid log(0)

# Add the response variable back in after transformations
ev_train_subset_log$total_ev_count <- ev_train$total_ev_count


lm_model_log <- lm(total_ev_count ~ ., data = ev_train_subset_log)


summary(lm_model_log)





```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.